---
title: "shake_genre_analysis"
author: "Ally Peyton"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Algorithm for Analysis

This markdown document contains all of the genre analysis - step 3 of my work - for this project. 

* Step 3: Genre Classification
In this analysis, I am approaching two tasks, each using three different kinds of classification: decision trees, naive bayes models, and svms. The first task is to classify two of Shakespeare's plays, The Tempest and Pericles, Prince of Tyre, based on word frequency. -
  + Step 3A: Preprocess, build, and tune naive Bayes model.
  + Step 3B: Preprocess, build, and tune SVM model.

### STEP 3: GENRE CLASSIFICATION
My first models will focus on classifying the genre of two plays, Pericles, Prince of Tyre, and the Tempest. 

### Step 3A: Attempting some Naive Bayes Modeling
Naive Bayes classification usually depends on data with categorical features, but the matricies are currently numeric. This next chunk of code will create a function to transform the matricies into categorical data with yes and no as the determiner for the presence of a word. 

First, though, I've discovered that somewhere within the categorization the names of the plays have gotten out of order. I think this has to do with some of the dplyr code from above, which I just don't feel comfortable trying to change, so I'm going to try to rebuild the dataset using the list data provided (above).

``` {r function for creating categorical matricies}
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

```

``` {r preparing data for NB models 2}
# create DTM from corpus of words
plays_dtm <- DocumentTermMatrix(plays_corpus_clean)

# find most frequent 1000 words
fterms <- findFreqTerms(plays_dtm, 1000)

# subset frequency terms from the dtm
plays_dtm_freq <- plays_dtm[, fterms]

# subset plays dtm into test and train dtms
genre_test_dtm <- plays_dtm_freq[c(26,31),]
genre_train_dtm <- plays_dtm_freq[-c(26,31),]

# move dtms into categorical for nb model
genre_test_cat <- apply(genre_test_dtm, MARGIN = 2, convert_counts)

genre_train_cat <- apply(genre_train_dtm, MARGIN = 2, convert_counts)

# create labels for nb model
genre_test_labels <- plays_genre[c(26,31),2]
genre_train_labels <- plays_genre[-c(26,31),2]

```

``` {r first naive bayes model}
nb1 <- naiveBayes(genre_train_cat, genre_train_labels, laplace = 0)

genre_pred1 <- predict(nb1, genre_test_cat, type = 'class')

#CrossTable(genre_pred1, genre_test_labels, prop.chisq = FALSE, prop.c = F, prop.r = F, dnn = c('predicted', 'actual'))

```

This model predicts that both the Tempest and Pericles are comedies. 

This isn't a very interesting model, though, so onto SVM modeling to see if there is anything more interesting there. 

### Step 3B: Attempting SVM Analysis
SVM analysis has to be performed using scaled numeric data - luckily, I have two matricies with numeric data, and the SVM model produced by the e1017 package does the scaling for you, so all I have to do is transform the document-term matricies into a dataframe with a target variable (eg. the sentiment or truth of each review). 

``` {r svm preprocessing}
# changing the matricies into dataframes
genre_test_df <- as.data.frame(as.matrix(genre_test_dtm))
genre_train_df <- as.data.frame(as.matrix(genre_train_dtm))

play_genre_labels <- plays$genre

genre_test_labels <- play_genre_labels[c(26, 31)]
genre_train_labels <- play_genre_labels[-c(26, 31)]

#genre_test_labels_df <- as.data.frame(genre_test_labels)
#genre_train_labels_df <- as.data.frame(genre_train_labels)

genre_test_df$id <- genre_test_labels
genre_train_df$id <- genre_train_labels
```

``` {r svm model 1}
genre_svm1 <- ksvm(id ~ . , data = genre_train_df, C = 1)
genre_svm_pred1 <- predict(genre_svm1, genre_test_df, type = 'response')
svm_accuracy1 <- genre_svm_pred1 == genre_test_df$id
table(svm_accuracy1)
#CrossTable(genre_svm_pred1, genre_test_df$id, prop.chisq = FALSE, prop.c = F, prop.r = F, dnn = c('predicted', 'actual'))

```

``` {r seeing if i can manage a decision tree }
# creating factor data to use with the DT
dt_train <- genre_train_df[-73]
dt_train$f_id <- as.factor(genre_train_labels)

dt_test <- genre_test_df[-73]
dt_test$f_id <- as.factor(genre_test_labels)

# creating the decision tree
dt1 <- C5.0(dt_train[-73], dt_train$f_id, trials = 1)
dtpred1 <- predict(dt1, dt_test, type = 'class')
summary(dt1)

```

This decision tree is actually pretty interesting. It uses the words "king", "one", "son", and "prince" to determine whether Shakespeare's plays are comedies, tragedies, history plays, or other. 

Pretty unsurprisingly, these three models all predicted that Pericles and the Tempest are comedies. I began with this analysis because it is simpler, and now I will move on to the next analysis: determining authorship of disputed Shakespearean works. 
